---
title: "Week 4 exercises"
author: "Wing-Ho Ko"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r}
library(tidyverse)
library(boot)
library(ggbiplot)
```

## Exercise 1: Writing your own function

Create a function that rescale the elements of vector on a linear scale from 0 to 1, where 0 is corresponds to the minimum value of the vector and 1 corresponds to the maximum value of the vector. Call your function `rescale01`


## Exercise 2: orca weights revisited

Return to the orca weights dataset (`orcas_simulated.csv`) from lecture 6. Construct the 95% percentile bootstrap interval for the difference in orca weight between 2024 and 2023


## Exercise 3: Salish Sea data revisited

Consider the water temperature data at Port Townsend. Again compare the mean from 2015 versus that from 2024. Construct the 95% BCa confidence interval of the difference in mean, ignoring the time dimension and treating the two dataset as unpaired


### Exercise 4: covariance and correlation in penguins dataset

Consider the 4 numeric variables in the penguins dataset: bill length, bill depth, flipper length, and body mass.

1. Compute the covariance matrix between these variables, after dropping the observations with missing measurements

2. Which pair of variables have the most positive covariance? Most negative?

3. Repeat (1) but with correlation matrix.

4. Repeat (2) but with correlations.


## Exercise 5: PCA on penguins dataset

Again consider the penguins dataset. Again consider the 4 numeric variables bill length, bill depth, flipper length, and body mass (retain also the `species` variable, for purpose that will be clear). Again first subset the data then drop any entries that have NA's

1. Perform principal component analysis on the dataset, and store the result in `penguins_pca`. Make sure you mean-subtracted and rescaled

2. Which variable has the strongest association with the first principal component? How about weakest association?

3. Make a cumulative scree plot (see the `type` argument of `ggscreeplot()`). If we want to keep enough principal components so that 80% of the correlation is explained, how many principal components do we need?

4. Produce two form biplot of the samples (first with PC1 and PC2, second with PC3 and PC4),and use color to distinguish between the three species

5. Based on the biplot(s), which principal component(s) are useful in separating the species, and which principal component(s) are not?
