---
title: "Bootstrap approach"
author: "Wing-Ho Ko"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r}
library(ggplot2)
library(tidyverse)
library(boot)
```


## Writing functions in R

### A simple example

A simple function that takes an input vector and create a mean-subtracted version of it

```{r}
mean_subtracted <- function(x){
  x - mean(x)
}
```

Apply the function on a vector and assign the return to a new vector

```{r}
x <- c(1, 8, 5, 7, 9)
y <- mean_subtracted(x)
print(y)
```

Note that the global `x` is not modified

```{r}
print(x)
```

### Explicit return

Make a function return early (before all lines are executed) using `return()`

```{r}
standardize_bad <- function(x){
  return(x - mean(x))
  x / sd(x) # useless line that never got executed
}
```

```{r}
standardize_bad(c(-2, 0, 2))
```

For the function to work properly, let it execute until the last line and returns the result

```{r}
standardize_good <- function(x){
  x - mean(x)
  x / sd(x)
}
```

```{r}
standardize_good(c(-2, 0, 2))
```

### Scoping and copy-on-write behaviors

A function cannot modify a global variable

```{r}
x <- 5  # global x
y <- -1 # global y

rewrite_xy <- function(){

  # a function that, alas, does nothing
  
  x <- 2  # local x
  y <- 10 # local y
}

rewrite_xy()

print(x)
print(y)
```

## Bootstrap confidence intervals

### One-sample example: Thornton Creek total phosphorus

The total phosphorus at various locations of water bodies are monitored by the King County. The data at the Thornton Creek mouth from 2015 on is given in the file `Thornton_phosphorus.csv`

Use bootstrap approach to construct the 95% confidence interval of mean total phosphorus concentration from the data, ignoring the time dimension

Load the file and inspect

```{r}
Thornton <- read_csv("data/Thornton_phosphorus.csv")
glimpse(Thornton)
```

Writing the function that calculate mean total phosphorus given a bootstrap sample

```{r}
mean_phosphorus <- function(data, index){
  data <- data[index,]
  mean(data$value)
}
```

Perform bootstrap sampling and inspect the output

```{r}
Thornton_boot <- boot(Thornton, mean_phosphorus, 2000)
Thornton_boot
```

Extract the mean value from the original sample and from the bootstrap samples

```{r}
Thornton_boot$t0
```

```{r}
car::some(Thornton_boot$t)
```

Find the percentile confidence interval

```{r}
Thronton_perc_ci <- boot.ci(Thornton_boot, 0.95, "perc")
Thronton_perc_ci
```

Extract the interval programmatically

```{r}
c(Thronton_perc_ci$percent[4], Thronton_perc_ci$percent[5])
```

Find the BCa confidence interval:

```{r}
Thronton_bca_ci <- boot.ci(Thornton_boot, 0.95, "bca")
Thronton_bca_ci
```

```{r}
print(Thronton_bca_ci)
```

Extract the interval programmatically

```{r}
c(Thronton_bca_ci$bca[4], Thronton_bca_ci$bca[5])
```

### Two-sample example: Thornton versus Tahlequah Creeks

Load the Tahlequah Creek data


```{r}
Tahlequah <- read_csv("data/Tahlequah_phosphorus.csv")
```

The two samples don't exactly line up, and for the purpose of this example we'll treat them as unpaired as opposed to paired

```{r}
c(nrow(Thornton), nrow(Tahlequah))
```

Step 1: combine the two data frames via `rbind()` ("row bind"). In addition, we'll code the "site" variable as factor in anticipation of using it in `boot()`

(factors in R are used to encode categorical variables that can take values from finitely many choices)

```{r}
phosphorus <- rbind(Thornton, Tahlequah) |> 
  mutate(site = as.factor(site))
```

Now we define the function that calculate the difference in means

```{r}
diff_mean <- function(data, index){
  data <- data[index,]
  Thornton <- filter(data, site=="Thornton Creek mouth at Sand Point Way NE")
  Tahlequah <- filter(data, site=="Tahlequah Creek mouth at SW Pohl Rd")
  mean(Thornton$value) - mean(Tahlequah$value)
}
```

Construct the bootstrap samples and perform the sampling

```{r}
phosphorus_boot <- boot(phosphorus, diff_mean, 5000, strata=phosphorus$site)
```

Now construct the BCa confidence interval:

```{r}
boot.ci(phosphorus_boot, 0.95, "bca")
```
