---
title: "covariance and correlation"
author: "Wing-Ho Ko"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r}
library(tidyverse)
```

## Covariance and correlation

### Example: noisy quintic curve

Loading the prepared dataset (quintic curve with noise)

```{r}
quintic <- read_csv("data/quintic_data.csv")
```

Plot the location of the data points

```{r}
ggplot(quintic) + geom_point(aes(x=x, y=y)) + coord_fixed()
```

Covariance between x and y

```{r}
cov(quintic$x, quintic$y)
```

Pearson's correlation between x and y

```{r}
cor(quintic$x, quintic$y)
```

Spearman's correlation between x and y

```{r}
cor(quintic$x, quintic$y, method="spearman")
```

Full covariance matrix

```{r}
cov(quintic)
```

Full correlation matrix

```{r}
cor(quintic)
```

### Example: bank notes

```{r}
banknotes <- read_csv("data/banknotes.csv")
glimpse(banknotes)
```

```{r}
banknotes_num <- select(banknotes, Length: Top)
```

Full covariance matrix

```{r}
cov(banknotes_num)
```

Full correlation matrix

```{r}
cor(banknotes_num)
```
