---
title: "Week 4 exercises comments"
author: "Wing-Ho Ko"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r}
library(tidyverse)
library(boot)
```

## Confidence intervals for paired samples

### Generate test data

```{r}
set.seed(123)

n <- 30

sample_1 <- rlnorm(n, 2, 1)
sample_2 <- sample_1 + rnorm(n, 5, 5)

pairs_long <- data.frame(
  tag = rep(1:n, times=2),
  year = rep(c(2020, 2021), each=n),
  value = c(sample_1, sample_2)
)
```

```{r}
pairs_long
```

### Checking assumption for normality

```{r}
ggplot(pairs_long, aes(sample=value)) + geom_qq() + geom_qq_line() + 
  facet_grid(~year)
```

```{r}
pairs_wide <- pivot_wider(
  pairs_long, 
  names_from="year", 
  names_prefix="year_",
  values_from="value"
) |> 
  mutate(diff = year_2021 - year_2020)

pairs_wide
```


```{r}
ggplot(pairs_wide, aes(sample=diff)) + geom_qq() + geom_qq_line()
```

Even though EACH sample is non-normal, it is actually fine to use t-test, since the **difference** is normally distributed!

### bootstrap confidence interval

```{r}
mean_diff <- function(data, indices){
  data <- data[indices,]
  mean(data$diff)
}
```

```{r}
pairs_boot <- boot(pairs_wide, mean_diff, 2000)
pairs_ci <- boot.ci(pairs_boot, type="bca")
pairs_ci
```


## Stratified vs non-stratified bootstrap

### Generating test data

Generate two samples from two uniform distributions

```{r}
set.seed(42)

n1 <- 100
n2 <- 10

sample_1 <- runif(n1, 1.5, 2.5)
sample_2 <- runif(n2, 0.25, 0.75)

test_data <- data.frame(
  label = c(rep(1, each=n1), rep(2, each=n2)),
  value = c(sample_1, sample_2)
)
```

```{r}
test_data
```


### Counting the number of observations drew from sample 2

A function that count the difference in number of observations from each sample in a bootstrap sample

```{r}
count_n1 <- function(data, index){
  data <- data[index,]
  sample_1 <- filter(data, label==1)
  nrow(sample_1)
}

count_n2 <- function(data, index){
  data <- data[index,]
  sample_2 <- filter(data, label==2)
  nrow(sample_2)
}
```

Stratified vs non-stratified bootstrap

```{r}
count_boot_ns <- boot(test_data, count_n2, 2000)
count_boot_s <- boot(test_data, count_n2, 2000, strata=test_data$label)
```

Non-stratified bootstrap include bootstrap samples for which the number of observations drew from sample 2 differ from the actual number of observations in sample 2.

```{r}
ggplot() + geom_bar(aes(x=count_boot_ns$t)) + xlim(-1, 25)
```

Stratified bootstrap only sample from the subspace where the number of observations drew from each sample is the same.

```{r}
ggplot() + geom_bar(aes(x=count_boot_s$t)) + xlim(-1, 25)
```

### Look at the difference in confidence interval

Function to calculate the difference in mean

```{r}
mean_ratio <- function(data, index){
  data <- data[index,]
  sample_1 <- filter(data, label==1)
  sample_2 <- filter(data, label==2)
  mean(sample_1$value) / mean(sample_2$value)
}
```

CI from non-stratified bootstrap

```{r}
ratio_boot_ns <- boot(test_data, mean_ratio, 5000)
boot.ci(ratio_boot_ns, type="bca")
```

CI from stratified bootstrap

```{r}
ratio_boot_s <- boot(test_data, mean_ratio, 5000, strata=test_data$label)
boot.ci(ratio_boot_s, type="bca")
```

Note that they are different beyond uncertainty introduced by finite R (= number of replicates). The stratified bootstrap gives the correct answer here because it corresponds to the sampling protocol of the original data

